Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-22 cav2017/cav2017.tex
text

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{CEGIS of Safe Controllers for LTI Systems with Solution Generalisation} 
\label{sec:CEGARIS} 
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%Abstractly-Refined
In this section we outline the general architecture of the progam synthesizer used 
in both approaches. We then detail the 2 methods of solution generalisation that we compare.
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{General architecture of the program synthesizer}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\label{synthesizer-general}
The input specification provided to the program synthesizer is of the form
$\exists \vec{P} .  \forall \vec{x}.  \sigma(\vec{x}, \vec{P})$ where
$\vec{P}$ ranges over functions, $\vec{x}$ \todo{now variables are vectorised, however previously the model variables aren't. please align notation. I do not like to use vectors unless essential, let's say it's a waste of ink in printouts. } ranges over ground terms and
$\sigma$ is a quantifier-free formula.  We interpret the ground terms over
some finite domain $\mathcal{D}$.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Synthesis problem: statement}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
The synthesis problem we are trying to solve is the following:
find a digital controller $K$ (see Eq.~\ref{eq:controlaction})
that makes the closed-loop system safe for 
initial state $x_0$, reference signal $r_k$ and input $u_k$
as defined in Sec.~\ref{ssec:safety}. 
We consider non-deterministic initial states within a specified range
and the reference signal to be 0 (\addtodo{refer to wherever this is explained initially}).
Then, when mapping back to the notation used for describing the general architecture 
of the program synthesizer, the controller $K$ denotes $P$ and 
$(x_0, u_k)$ represents $x$. 
While the controller's precision $\langle I_c,F_c\rangle$ is given, 
we can vary the plant's precision $\langle I_p,F_p\rangle$ such that 
$\mathbb{R}\langle I_p,F_p \rangle \supseteq \mathbb{R}\langle I_c,F_c \rangle$.

@@

We investigate two different approaches of integrating solution generalisation 
with CEGIS:
\begin{itemize}
\item[1.] The first approach described in Sec.~\ref{sec:CEGIS-precision-incrementation}
  uses a multi-staged validation phase and performs solution generalisation based on
the plant's precision, the number of loop unrollings
and the sampling rate. Essentially, it finds a controller for an individual initial state
and input with a bounded time horizon and generalizes to all states,
inputs and time horizons. %% (with a final verification loop that verifies
%% the continuous system behaviour against the previously verified
%% discrete system bi-simulation)
\item[2.] The second approach described in Sec~\ref{sec:CEGIS-abstraction-refinement}
  %% uses abstract acceleration to compute the set of reachable states,
  %% which enables it to perform abstraction refinement inside the CEGIS loop.
  finds a controller for a continuous initial
state-space and input sets over an abstraction of the continuous
dynamics which conforms to witness proofs at specific times, and
generalizes the solution for all times by refining the number of
witnesses.
\end{itemize}

@@

 \end{tikzpicture}
}}
\caption{CEGIS with multi-staged verification.}
\label{fig:CEGIS-precision-increment}
\end{figure}

@@

The controller synthesis is described in Fig.~\ref{fig:CEGIS-precision-increment}.
One important observation is that, for this approach, we verify and synthesise 
a controller over $k$ time steps. We then compute a completeness threshold, $\overline{k}$,
for this controller, and verify the controller's correctness to $\overline{k}$ time steps.
Essentially, $\overline{k}$ represents the number of iterations 
required to sufficiently unwind the closed-loop state-space system 
such that the boundaries are not violated for any $k{>}\overline{k}$
(i.e. it's sufficient to unwound the closed-loop state-space system up to  
$\overline{k}$ in order to ensure that $\phi_{safety}$ holds).

@@

@@

@@

>>>>

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{CEGIS of Safe Controllers for LTI Systems with Solution Generalisation} 
\label{sec:CEGARIS} 
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%Abstractly-Refined
In this section we outline the general architecture of the program synthesiser used in both approaches\todo{which approaches?}. 
We then detail the two methods of solution generalisation that we compare.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{General architecture of the program synthesiser}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\label{synthesizer-general}
The input specification provided to the program synthesizer is of the form
$\exists \vec{P} .  \forall \vec{x}.  \sigma(\vec{x}, \vec{P})$ where
$\vec{P}$ ranges over functions, $\vec{x}$ \todo{now variables are vectorised, however previously the model variables aren't. please align notation. I do not like to use vectors unless essential, let's say it's a waste of ink in printouts. } ranges over ground terms and
$\sigma$ is a quantifier-free formula.  We interpret the ground terms over
some finite domain $\mathcal{D}$.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Synthesis problem: statement (recap) and connection to program synthesis}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

\todo{The following paragraph has been already stated above. Let us clarify this is a useful repetition to recap things?}

The synthesis problem we are trying to solve is the following:
find a digital controller $K$ (see Eq.~\ref{eq:controlaction})
that makes the closed-loop system safe for 
initial state $x_0$, reference signal $r_k$ and input $u_k$
as defined in Sec.~\ref{ssec:safety}. 
We consider non-deterministic initial states within a specified range
and the reference signal to be set to zero.

When mapping back to the notation used for describing the general architecture 
of the program synthesiser, the controller $K$ denotes $P$ and 
$(x_0, u_k)$ represents $x$. 
While the controller's precision $\langle I_c,F_c\rangle$ is given, 
we can vary the plant's precision $\langle I_p,F_p\rangle$ such that 
$\mathbb{R}\langle I_p,F_p \rangle \supseteq \mathbb{R}\langle I_c,F_c \rangle$.

@@

We investigate two different approaches of integrating solution generalisation 
with CEGIS:
\begin{itemize}
\item[1.] The first approach (described in Sec.~\ref{sec:CEGIS-precision-incrementation})
  uses a multi-staged verification phase and performs solution generalisation based on
the plant precision, the number of loop unrollings, 
and the sampling rate. Essentially, it finds a controller for an individual initial state
and input with a bounded time horizon and generalises to all states,
inputs, and time horizons. %% (with a final verification loop that verifies
%% the continuous system behaviour against the previously verified
%% discrete system bi-simulation)
\item[2.] The second approach (described in Sec~\ref{sec:CEGIS-abstraction-refinement})
  %% uses abstract acceleration to compute the set of reachable states,
  %% which enables it to perform abstraction refinement inside the CEGIS loop.
  finds a controller for a continuous initial set of states and input set, 
  over an abstraction of the continuous dynamics that conforms to witness proofs at specific times; 
  it further generalises the solution over all times by refining the number of witnesses. 
\end{itemize}

@@

 \end{tikzpicture}
}}
\caption{CEGIS with multi-staged verification}
\label{fig:CEGIS-precision-increment}
\end{figure}

@@

The controller synthesis is described in Fig.~\ref{fig:CEGIS-precision-increment}.
One important observation is that, for this approach, we verify and synthesise 
a controller over $k$ time steps. We then compute a completeness threshold, $\overline{k}$,
for this controller, and verify the controller correctness to $\overline{k}$ time steps.
Essentially, $\overline{k}$ represents the number of iterations 
required to sufficiently unwind the closed-loop state-space system 
such that the boundaries are not violated for any $k{>}\overline{k}$
(i.e. it is sufficient to unwind the closed-loop state-space system up to  
$\overline{k}$ in order to ensure that $\phi_{safety}$ holds).

\medskip

@@

\todo{please eventually itemize next}

@@

Proof. 
Consider the set of initial states, $X_0$, assumed to be convex since it's a box. 
Name $v_i$ its vertices, where $i=1,\ldots, 2^n$.  
Thus any point $x \in X_0$ can be expressed by convexity as $x = \sum_{i=1}^{2^n} \alpha_i v_i$, 
where $\sum_{i=1}^{2^n} \alpha_i =1$. 
So then if $x_0=x$, then $x_k = (A_d - B_d K)^k x = 
(A_d - B_d K)^k \sum_{i=1}^{2^n} \alpha_i v_i = 
\sum_{i=1}^{2^n} \alpha_i (A_d - B_d K)^k v_i  = 
\sum_{i=1}^{2^n} \alpha_i x_k^i$, 
where I have denoted with $x_k^i$ the trajectories obtained from the single vertices $v_i$.QED

@@

Proof sketch. 
Consider spectrum of matrix $(A_d - B_d K)$ and select eigenvalue with smallest (non trivial) imaginary value.  
call such value $\vartheta$. 
At every time interval $kT_s, (k+1)T_s$ ($T_s$ being the sampling time), 
the dynamics projected on the corresponding eigenspace rotate of $\vartheta T_s$ radiants.  
So surely taking $k$ as the ceiling of $\frac{2\pi}{\vartheta T_s}$, after $k$ steps we've completed a rotation. 
This is the CT.
